---
title: "Explore VCF"
author: "AL"
date: "29/03/2021"
output:
  html_document:
    toc: true
    number_sections: true
    toc_float: true
editor_options:
  chunk_output_type: console
---

# Summary

- An interesting pattern in **Qual**: with some multimodality  
- Some mitochondrial variants have no data in **Qual**  
- Apparently, there is no * in **Alt**  
- There is no **ID**-s and all variants **PASS**-ed filters  

<style>
pre{
  overflow-x: auto;
}
pre code{
  word-wrap: normal;
  white-space: pre;
}
s</style>

```{r echo=F}
options(width=999)
```

# Start section

```{r}

Sys.time()
rm(list=ls())
graphics.off()
gc()

```

# Read data

reading may take some minutes in m4.large

```{r}

data_folder <- "/home/share/data/s03_split_multiallelic_sites"
data_file <- "IHCAPX8_dragen_joint.hard-filtered.PF.MA-flag.MA-split.annotations.txt"
data.df <- read.table(file.path(data_folder,data_file),
                      header=F,sep="\t",quote="")

head(data.df)
#View(head(data.df))

colnames(data.df) <- c("chr","pos","id","ref","alt","filter","qual","info")

dim(data.df)
str(data.df)

rm(data_file)

```

# Qual

```{r}

x <- data.df$qual
x <- as.numeric(x)
sum(is.na(x))
data.df$qual[is.na(x)]

#View(data.df[is.na(x),])
data.df[is.na(x),]

data.df$qual <- as.numeric(data.df$qual)
hist(data.df$qual, main="QUAL", xlab="qual")
hist(data.df$qual[data.df$qual<600], main="QUAL (<600)", xlab="qual")
quantile(data.df$qual, na.rm=T)

rm(x)

```

# No star in ALt ?

```{r}

x <- grepl("\\*",data.df$alt)
sum(x)
rm(x)

```

# Multiallelic sites

```{r}

x <- grepl("MULTIALLELIC",data.df$info)
sum(x)

#View(head(data.df[x,]))
head(data.df[x,])

rm(x)

```

# Filter and Id

```{r}
table(data.df$id, useNA = "always")
table(data.df$filter, useNA = "always")
```

# Save data

```{r}

save(data.df,file=file.path(data_folder,"IHCAPX8_dragen_joint.hard-filtered.PF.MA-flag.MA-split.annotations.RData"))

```

# Final section

```{r}
ls()
gc()
Sys.time()
sessionInfo()
```
